---
title: "311 EDA"
author: "Ana Paula Felix de Queiroz"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{=html}
<style type="text/css">

h1.title {
  font-size: 38px;
  font-family: "Times New Roman", Times, serif;
  color: #2a6592;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: #2a6592;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(car)

```


## 1. Load the Dataset In R



```{r}
# Set the working directory to the folder containing the CSV file
setwd("C:/Users/pauli/OneDrive/Área de Trabalho/git_repo/DS311-EDA/EDA_Python/data")

# Read the CSV file into a data frame
ames <- read.csv("ames.csv")

# Verify if the data was loaded correctly
head(ames)
```


Inspect the contents of the dataframe:

```{r}


str(ames)
dim(ames)
```


## 2. Explore Data Distributions

Write code to produce histograms showing the distributions of `SalePrice`, `TotRmsAbvGrd`, and `OverallCond`.



### Sale Price

In the cell below, produce a histogram for `SalePrice`.

```{r}


df = ames

# Extract the relevant data
data <- df$SalePrice
mean <- mean(data)




# Set up plot
hist(data, main = "Distribution of Sale Prices", xlab = "Sale Price", ylab = "Number of Houses", col = "lightblue", border = "black", breaks = 80)

# Add vertical line
abline(v = mean, col = "black", lwd = 2)
```



Now, print out the mean, median, and standard deviation:

```{r}

mean = mean(data)

median = median(data)

 sd = sd(data)
 
 print(paste("Mean:", mean))
  print(paste("Median:", median))
print(paste("Standard Deviation:", sd))
```
### Total Rooms Above Grade

In the cell below, produce a histogram for `TotRmsAbvGrd`.

```{r}

# Extract the relevant data
data <- df$TotRmsAbvGrd

mean <- mean(data)

# Set up plot
hist(data, main = "Distribution of Sale Prices", xlab = "Sale Price", ylab = "Number of Houses", col = "lightblue", border = "black", breaks = 80)

# Add vertical line
abline(v = mean, col = "black", lwd = 2)
```


Now, print out the mean, median, and standard deviation:

```{r}

mean = mean(data)

median = median(data)

 sd = sd(data)
 
 print(paste("Mean:", mean))
  print(paste("Median:", median))
print(paste("Standard Deviation:", sd))


```

### Overall Condition

In the cell below, produce a histogram for `OverallCond`.



```{r}

# Extract the relevant data
data <- df$OverallCond
mean <- mean(data)

# Set up plot
hist(data, main = "Distribution of Sale Prices", xlab = "Sale Price", ylab = "Number of Houses", col = "lightblue", border = "black", breaks = 80)

# Add vertical line
abline(v = mean, col = "black", lwd = 2)
```

Now, print out the mean, median, and standard deviation:
```{r}

mean = mean(data)

median = median(data)

 sd = sd(data)
 
 
 print(paste("Mean:", mean))
  print(paste("Median:", median))
print(paste("Standard Deviation:", sd))



```

## 3. Explore Differences between Subsets

As you might have noted in the previous step, the overall condition of the house seems like we should treat it as more of a categorical variable, rather than a numeric variable.

One useful way to explore a categorical variable is to create subsets of the full dataset based on that categorical variable, then plot their distributions based on some other variable. Since this dataset is traditionally used for predicting the sale price of a house, let's use `SalePrice` as that other variable.

In the cell below, create three variables, each of which represents a record-wise subset of `df` (meaning, it has the same columns as `df`, but only some of the rows).

-   `below_average_condition`: home sales where the overall condition was less than 5
-   `average_condition`: home sales where the overall condition was exactly 5
-   `above_average_condition`: home sales where the overall condition was greater than 5

```{r}
below_average_condition <- df[df$OverallCond < 5, ]
above_average_condition <- df[df$OverallCond > 5, ]
average_condition <- df[df$OverallCond == 5, ]

library(scales)


# Calculate the range of SalePrice
price_range <- range(c(above_average_condition$SalePrice, average_condition$SalePrice, below_average_condition$SalePrice))

# Create custom bins
bins <- seq(price_range[1], price_range[2], length.out = 80)  # Adjust the length.out as needed

# Set up plot
par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
plot.new()
plot.window(xlim = price_range, ylim = c(0, 70), xlab = "Sale Price", ylab = "Number of Houses")
title(main = "Distributions of Sale Price Grouped by Condition")

# Plot histograms
hist(above_average_condition$SalePrice, breaks = bins, col = "cyan", alpha = 0.5, add = TRUE)

hist(average_condition$SalePrice, breaks = bins, col = "gray", alpha = 0.3, add = TRUE)
hist(below_average_condition$SalePrice, breaks = bins, col = "yellow", alpha = 0.5, add = TRUE)

# Add legend with transparency
legend("topright", legend = c("above average condition", "average condition", "below average condition"),
       fill = c(alpha("cyan", 0.5), alpha("gray", 0.3), alpha("yellow", 0.5)), density = c(0, 0, 0))

# Customize labels
axis(1)
axis(2)


```


## 4. Explore Correlations

To understand more about what features of these homes lead to higher sale prices, let's look at some correlations. We'll return to using the full `df`, rather than the subsets.

In the cell below, print out both the name of the column and the Pearson correlation for the column that is ***most positively correlated*** with `SalePrice` (other than `SalePrice`, which is perfectly correlated with itself).

We'll only check the correlations with some kind of numeric data type.

You can import additional libraries, although it is possible to do this just using pandas.



```{r}
# Load required libraries
library(dplyr)
library(Hmisc)

# Create a data frame that contains only numeric columns
numeric_df <- df %>% select_if(is.numeric) %>% select(-SalePrice, -Id)

correlations <- vector("numeric")

for (col in colnames(numeric_df)) {
  corr <- rcorr(df[[col]], df$SalePrice)$r[1, 2]
  correlations <- c(correlations, corr)
}

max_corr_value <- max(correlations)
max_corr_column <- colnames(numeric_df)[which(correlations == max_corr_value)]

cat("Most Positively Correlated Column:", max_corr_column, "\n")
cat("Maximum Correlation Value:", max_corr_value, "\n")




min_corr_value <- min(correlations)
min_corr_column <- colnames(numeric_df)[which(correlations == min_corr_value)]

cat("Most Negativelly Correlated Column:", min_corr_column, "\n")
cat("Minimun Correlation Value:", min_corr_value, "\n")



```

Once you have your answer, edit the code below so that it produces a scatter plot of the relevant columns.

```{r}


library(ggplot2)
library(gridExtra)

# Convert x-axis variables to factors
df$OverallQual <- as.factor(df$OverallQual)
df$KitchenAbvGr <- as.factor(df$KitchenAbvGr)

# Define color palette
colors <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
            "#D55E00", "#CC79A7", "#999999", "#009999", "#66CCFF")

# Plot distribution of column with highest correlation
p1 <- ggplot(df, aes(x = OverallQual, y = SalePrice, fill = OverallQual)) +
  geom_boxplot() +
  scale_fill_manual(values = colors, breaks = levels(df$OverallQual)) +
  labs(title = "Overall Quality vs. Sale Price",
       x = "Overall Quality",
       y = "Sale Price")

# Plot distribution of column with most negative correlation
p2 <- ggplot(df, aes(x = KitchenAbvGr, y = SalePrice, fill = KitchenAbvGr)) +
  geom_boxplot() +
  scale_fill_manual(values = colors, breaks = levels(df$KitchenAbvGr)) +
  labs(title = "Number of Kitchens vs. Sale Price",
       x = "Number of Kitchens Above Ground",
       y = "Sale Price")

# Arrange the plots side by side
grid.arrange(p1, p2, ncol = 2)

```


## 5. Engineer and Explore a New Feature

Here the code is written for you, all you need to do is interpret it.

We note that the data spans across several years of sales:


```{r}
table(df$YrSold)

```

Maybe we can learn something interesting from the age of the home when it was sold. This uses information from the YrBuilt and YrSold columns, but represents a truly distinct feature.


```{r}
# Make a new column, Age
df$Age <- df$YrSold - df$YearBuilt

# Set up plot
library(ggplot2)
ggplot(df, aes(x = Age, y = SalePrice)) +
  geom_point(alpha = 0.3, color = "darkgreen") +
  labs(title = "Home Age vs. Sale Price",
       x = "Age of Home at Time of Sale",
       y = "Sale Price")

```

